Human capital mediates natural selection in contemporary humans

### 2021-12-30 why-natural-selection.Rmd
tweak intro

As yet we know little about the social mechanisms behind these effects. A
potential explanation for them comes from the economic theory of fertility first
presented by [@becker1960economic]. In this theory, higher potential earnings
have two opposite effects on fertility: a fertility-increasing *income effect*
(higher income makes children more affordable), and a fertility-lowering
*substitution effect* (time spent on childrearing has a higher cost in foregone
earnings). Thus, an individual's *human capital* -- skills and personality
traits which are valuable in labour markets -- can increase or decrease his or
her fertility. This can lead to selection on inherited
variants which are linked to human capital. Furthermore, the economic theory predicts
systematic variation in the relative strength of the income and substitution
effects across different social groups.

@@

We then show that a parsimonious model of human capital, education and fertility
choices can give rise to these empirical results. To test the theory, we show
that polygenic scores' correlation with fertility is associated with their
correlation with education and earnings. Furthermore, controlling for education
and earnings reduces the correlation with fertility, suggesting that human
capital mediates natural selection. Thus, contemporary natural selection on
polygenic scores can be explained by their correlation with earnings-increasing
human capital.

>>>>

As yet we know little about the social mechanisms behind natural selection. A
potential explanation comes from the economic theory of fertility [@becker1960economic]. In this theory, higher potential earnings
have two opposite effects on fertility: a fertility-increasing *income effect*
(higher income makes children more affordable), and a fertility-lowering
*substitution effect* (time spent on childrearing has a higher cost in foregone
earnings). Thus, an individual's *human capital* -- skills and personality
traits which are valuable in labour markets -- can increase or decrease his or
her fertility. This can lead to selection on inherited
variants which are linked to human capital. Furthermore, the economic theory predicts
systematic variation in the relative strength of the income and substitution
effects across different social groups.

@@

We then show that a simple model of human capital, education and fertility
choices can give rise to these empirical results. To test the theory, we show
that polygenic scores' correlation with fertility is associated with their
correlation with education and earnings. Furthermore, a mediation analysis shows
that part of the correlation with fertility is indeed *via* education. Thus, contemporary natural selection on polygenic scores can be explained by their correlation with earnings-increasing human capital.
